Titeln på avhandlingens språk

### 2023-05-19 body/part1/chapter4-conclusions/conclusions.tex
fixes

Finally, I see tremendous opportunities for testbed-driven research in Edge Computing.
My current efforts in this context aim towards the extension of our Ainur framework and its integration with \gls{CHI}~\cite{keahey2020lessons}.
I believe Ainur's unique focus on fully-automated end-to-end wireless Edge Computing experimentation makes it valuable addition to the \gls{CHI} stack.
At the same time, integrating with \gls{CHI} will provide a number of features and functionalities which will greatly expand Ainurs potential, such as resource reservation, multi-tenant testbed access, and automated bare-metal and \gls{VM}-host instantiation.

>>>>

Finally, I see tremendous opportunities for testbed-driven research in Edge Computing.
My current efforts in this context aim towards the extension of our Ainur framework and its integration with \gls{CHI}~\cite{keahey2020lessons}.
I believe Ainur's unique focus on fully-automated end-to-end wireless Edge Computing experimentation makes it valuable addition to the \gls{CHI} stack.
At the same time, integrating with \gls{CHI} will provide a number of features and functionalities which will greatly expand Ainur's potential, such as resource reservation, multi-tenant testbed access, and automated bare-metal and \gls{VM}-host instantiation.

### 2023-05-19 publications/2023EdgeDroid2/abstract.tex
fixes

    We present in this paper the first model for the complete end-to-end emulation of humans in \gls{WCA}.
    We build this model through statistical analysis of data collected from previous work in this field, and demonstrate its utility by studying application task durations.
    Compared to first-order approximations, our model shows a \textasciitilde\SI{36}{\percent} larger gap between step execution times at high system impairment versus low.
    We further introduce a novel framework for stochastic optimization of resource consumption-responsiveness tradeoffs in \gls{WCA}, and show that by combining this framework with our realistic model of human behavior, significant reductions of up to \SI{50}{\percent} in number processed frame samples and \SI{20}{\percent} in energy consumption can be achieved with respect to the state-of-the-art.
\end{abstract}
\glsresetall%

>>>>

    We present in this paper the first model for the complete end-to-end emulation of humans in \gls{WCA}.
    We build this model through statistical analysis of data collected from previous work in this field, and demonstrate its utility by studying application task durations.
    Compared to first-order approximations, our model shows a \ensuremath{\sim}\SI{36}{\percent} larger gap between step execution times at high system impairment versus low.
    We further introduce a novel framework for stochastic optimization of resource consumption-responsiveness tradeoffs in \gls{WCA}, and show that by combining this framework with our realistic model of human behavior, significant reductions of up to \SI{50}{\percent} in number processed frame samples and \SI{20}{\percent} in energy consumption can be achieved with respect to the state-of-the-art.
\end{abstract}
\glsresetall%

### 2023-05-19 publications/2023EdgeDroid2/implications_footprint.tex
fixes

The results of this investigation are presented in \cref{fig:lifetimes}.
Compared to the reference model, our realistic model is, on average, roughly \SI{11}{\percent} faster when subject to low \glspl{TTF}.
At higher \glspl{TTF}, the behavior of the model depends on its level of normalized neuroticism.
At low neuroticism, the behavior of the realistic model results in total task durations that are basically indistinguishable from the reference model.
However, at high neuroticism, the model once again results in a considerable difference in total task duration with respect to the reference --- this time extending durations by \textasciitilde\SI{11}{\percent} on average.

@@

We deploy configurations running \num{45}-step versions of the LEGO task described in~\cite{olguinmunoz2021impact}.
The testbed configurations include setups with \numlist{1;4;7;10} clients.
Due to having limited time, each combination of testbed and timing model configuration is only repeated \num{10} times.
The results are presented in \cref{fig:testbed_lifetimes}.
Owing to the low number of samples, specifically to minimize the effects of potential outliers, we opt here for the geometric rather than arithmetic mean to represent our results.
The results are nonetheless clear, and follow the same pattern as the previously discussed results under ideal, controlled circumstances.
With just a single client and mean \gls{RTT} of around \SI{400}{\milli\second}, all parameterizations of the realistic model achieved an average task duration \SIrange{6}{12}{\percent} shorter than the reference.
At \num{10} clients, the results mimic those at higher \glspl{TTF} in the ideal controlled setup, with high neuroticism parameterizations reaching \textasciitilde\SI{8}{\percent} longer application lifetimes.

>>>>

The results of this investigation are presented in \cref{fig:lifetimes}.
Compared to the reference model, our realistic model is, on average, roughly \SI{11}{\percent} faster when subject to low \glspl{TTF}.
At higher \glspl{TTF}, the behavior of the model depends on its level of normalized neuroticism.
At low neuroticism, the behavior of the realistic model results in total task durations that are basically indistinguishable from the reference model.
However, at high neuroticism, the model once again results in a considerable difference in total task duration with respect to the reference --- this time extending durations by \ensuremath{\sim}\SI{11}{\percent} on average.

@@

We deploy configurations running \num{45}-step versions of the LEGO task described in~\cite{olguinmunoz2021impact}.
The testbed configurations include setups with \numlist{1;4;7;10} clients.
Due to having limited time, each combination of testbed and timing model configuration is only repeated \num{10} times.
The results are presented in \cref{fig:testbed_lifetimes}.
Owing to the low number of samples, specifically to minimize the effects of potential outliers, we opt here for the geometric rather than arithmetic mean to represent our results.
The results are nonetheless clear, and follow the same pattern as the previously discussed results under ideal, controlled circumstances.
With just a single client and mean \gls{RTT} of around \SI{400}{\milli\second}, all parameterizations of the realistic model achieved an average task duration \SIrange{6}{12}{\percent} shorter than the reference.
At \num{10} clients, the results mimic those at higher \glspl{TTF} in the ideal controlled setup, with high neuroticism parameterizations reaching \ensuremath{\sim}\SI{8}{\percent} longer application lifetimes.

### 2023-05-19 publications/2023EdgeDroid2/intro.tex
mark as prelim
fixes

\todo[inline]{Needs rewrite}

@@

In that context, the contributions of this paper are:
\begin{enumerate}
    \item\label{item:contrib:model} We introduce the first, to our knowledge, stochastic model for human timings in \gls{WCA} applications.
    Using the data collected for~\cite{olguinmunoz2021impact} as a base, we build a stochastic model which takes as input past measurements of system responsiveness and produces realistic step execution times.
    We also introduce a novel way to generate dynamic traces of frames for \gls{WCA} applications which can be combined with the timing model for a full end-to-end emulation of a human.
    We name this new model \emph{\edgedroid}; a direct, more realistic evolution of our initial EdgeDroid approach~\cite{olguinmunoz2018demoscaling,olguinmunoz2019edgedroid}.
    \item\label{item:contrib:footprint} Using this model, we study the implications of realistic human behavior for the \emph{application lifetime footprint} of \gls{WCA}, understanding this term as the duration of a specific complete execution of the application task.
    In accordance with previous work~\cite{olguinmunoz2021impact}, we find dependencies between system responsiveness and human step execution times that lead to substantially different application lifetimes when compared to a first-order baseline which does not take into account human behavior.
    \item\label{item:contrib:optimization} Finally, we study the potential for optimization in \gls{WCA} when considering human behavior using our model.
    We develop a generic model for the stochastic optimization of resource consumption versus responsiveness trade-offs in these applications, which we apply to two potential avenues for \gls{WCA} optimization; number of processed samples and energy consumption per step.
    First, we study the potential for reducing the number of samples captured per step.
    This is a valuable endeavor, as reducing the number of samples captured and subsequently processed directly translates in lower bandwidth demand on the wireless network and processor time demand on the cloudlet.
    Next, we explore the potential for direct optimization of energy consumption.
    The economic feasibility of WCA, and hence its likelihood of commercial adoption depends on it not being a resource hog, and this work furthers that effort.
    Leveraging our more involved user model, combined with the introduced optimization approach, we achieve a \textasciitilde\SI{60}{\percent} reduction in the number of samples processed per step.
    We also achieve an improvement of \SI{20}{\percent} in energy consumption, all while maintaining comparable levels of system responsiveness.
\end{enumerate}

>>>>

@@

In that context, the contributions of this paper are:
\begin{enumerate}
    \item\label{item:contrib:model} We introduce the first, to our knowledge, stochastic model for human timings in \gls{WCA} applications.
    Using the data collected for~\cite{olguinmunoz2021impact} as a base, we build a stochastic model which takes as input past measurements of system responsiveness and produces realistic step execution times.
    We also introduce a novel way to generate dynamic traces of frames for \gls{WCA} applications which can be combined with the timing model for a full end-to-end emulation of a human.
    We name this new model \emph{\edgedroid}; a direct, more realistic evolution of our initial EdgeDroid approach~\cite{olguinmunoz2018demoscaling,olguinmunoz2019edgedroid}.
    \item\label{item:contrib:footprint} Using this model, we study the implications of realistic human behavior for the \emph{application lifetime footprint} of \gls{WCA}, understanding this term as the duration of a specific complete execution of the application task.
    In accordance with previous work~\cite{olguinmunoz2021impact}, we find dependencies between system responsiveness and human step execution times that lead to substantially different application lifetimes when compared to a first-order baseline which does not take into account human behavior.
    \item\label{item:contrib:optimization} Finally, we study the potential for optimization in \gls{WCA} when considering human behavior using our model.
    We develop a generic model for the stochastic optimization of resource consumption versus responsiveness trade-offs in these applications, which we apply to two potential avenues for \gls{WCA} optimization; number of processed samples and energy consumption per step.
    First, we study the potential for reducing the number of samples captured per step.
    This is a valuable endeavor, as reducing the number of samples captured and subsequently processed directly translates in lower bandwidth demand on the wireless network and processor time demand on the cloudlet.
    Next, we explore the potential for direct optimization of energy consumption.
    The economic feasibility of WCA, and hence its likelihood of commercial adoption depends on it not being a resource hog, and this work furthers that effort.
    Leveraging our more involved user model, combined with the introduced optimization approach, we achieve a \ensuremath{\sim}\SI{60}{\percent} reduction in the number of samples processed per step.
    We also achieve an improvement of \SI{20}{\percent} in energy consumption, all while maintaining comparable levels of system responsiveness.
\end{enumerate}

### 2023-05-19 publications/2023EdgeDroid2/model.tex
fixes

We begin by calculating rolling weighted averages of the \glspl{TTF} of each of the \num{40} individual repetitions of the data.
We use exponentially decaying weights for the most recent \num{12} steps, defined in \cref{eq:weights}, such that the most recent \gls{TTF} accounts for roughly \SI{50}{\percent} of the rolling average, the second-most recent for \textasciitilde\SI{25}{\percent}, the third-most \textasciitilde\SI{12}{\percent}, and so on.
We additionally pad the data for each run with \num{12} copies of the first \gls{TTF} in order to ensure sensible values for the first twelve steps; this padding is removed after the weighted values have been calculated.
This weighing ensures that drastic changes in \gls{TTF} are significantly remembered by the model for at least \num{4} steps --- in line with our previous findings on human behavior --- and are subsequently quickly forgotten.
\begin{equation}\label{eq:weights}
    w_{n - i} = 
    \left\{ \begin{array}{ll}
        \frac{e^{-0.7 i}}{\sum\limits^{12}_{j=1} e^{-0.7 j}} & 1 \leq i \leq 12 \\
        & \\
        0 & i > 12
    \end{array} \right.
\end{equation}

@@

\cref{fig:ttf_to_exectime} shows the mean execution time outputted by the model when fed three different levels of \gls{TTF} (\emph{low}, \SI{0}{\second}, \emph{medium}, \SI{2.5}{\second}, or \emph{high}, \SI{5}{\second}).
These results were generated by first warming up the model by feeding it \num{25} \glspl{TTF} selected at random from the data before feeding it the desired input \gls{TTF}, and recording the generated execution time.
This procedure is repeated \num{600} times for each configuration and target \gls{TTF}.
The resulting mean execution times match precisely the desired behavior mentioned in \cref{it:ttftoexectime}, with the difference in mean execution times at low versus high \glspl{TTF} reaching \SI{14}{\percent} (\textasciitilde\SI{5.6}{\second} to \textasciitilde\SI{6.4}{\second}) in the worst case (high neuroticism configuration).
Additionally, we can observe the effects of neuroticism on generated execution times, as specified in \cref{it:neuro}.
At low neuroticism, the average difference between execution times at low versus high \glspl{TTF} was of roughly \SI{6.2}{\percent}, compared to \SI{12.5}{\percent} at high neuroticism.

@@

\begin{figure}
    \centering
    \includegraphics[width=.9\textwidth]{publications/2023EdgeDroid2/model_data/frame_probabilities}
    \caption{%
        Probability density of frame result tags as a step progresses.
        Note that \texttt{SUCCESS} frames are not included as --- by definition --- the probability for success frames is \num{1} for all normalized instant values greater than or equal to \num{1.0}.
    }\label{fig:frameprobs}
\end{figure}

>>>>

We begin by calculating rolling weighted averages of the \glspl{TTF} of each of the \num{40} individual repetitions of the data.
We use exponentially decaying weights for the most recent \num{12} steps, defined in \cref{eq:weights}, such that the most recent \gls{TTF} accounts for roughly \SI{50}{\percent} of the rolling average, the second-most recent for \ensuremath{\sim}\SI{25}{\percent}, the third-most \ensuremath{\sim}\SI{12}{\percent}, and so on.
We additionally pad the data for each run with \num{12} copies of the first \gls{TTF} in order to ensure sensible values for the first twelve steps; this padding is removed after the weighted values have been calculated.
This weighing ensures that drastic changes in \gls{TTF} are significantly remembered by the model for at least \num{4} steps --- in line with our previous findings on human behavior --- and are subsequently quickly forgotten.
\begin{equation}\label{eq:weights}
    w_{n - i} = 
    \left\{ \begin{array}{ll}
        \frac{e^{-0.7 i}}{\sum\limits^{12}_{j=1} e^{-0.7 j}} & 1 \leq i \leq 12 \\
        & \\
        0 & i > 12
    \end{array} \right.
\end{equation}

@@

\cref{fig:ttf_to_exectime} shows the mean execution time outputted by the model when fed three different levels of \gls{TTF} (\emph{low}, \SI{0}{\second}, \emph{medium}, \SI{2.5}{\second}, or \emph{high}, \SI{5}{\second}).
These results were generated by first warming up the model by feeding it \num{25} \glspl{TTF} selected at random from the data before feeding it the desired input \gls{TTF}, and recording the generated execution time.
This procedure is repeated \num{600} times for each configuration and target \gls{TTF}.
The resulting mean execution times match precisely the desired behavior mentioned in \cref{it:ttftoexectime}, with the difference in mean execution times at low versus high \glspl{TTF} reaching \SI{14}{\percent} (\ensuremath{\sim}\SI{5.6}{\second} to \ensuremath{\sim}\SI{6.4}{\second}) in the worst case (high neuroticism configuration).
Additionally, we can observe the effects of neuroticism on generated execution times, as specified in \cref{it:neuro}.
At low neuroticism, the average difference between execution times at low versus high \glspl{TTF} was of roughly \SI{6.2}{\percent}, compared to \SI{12.5}{\percent} at high neuroticism.

@@

\begin{figure}
    \centering
    \includegraphics[width=.9\textwidth]{model_data/frame_probabilities}
    \caption{%
        Probability density of frame result tags as a step progresses.
        Note that \texttt{SUCCESS} frames are not included as --- by definition --- the probability for success frames is \num{1} for all normalized instant values greater than or equal to \num{1.0}.
    }\label{fig:frameprobs}
\end{figure}

### 2023-05-19 publications/2023EdgeDroid2/old_model.tex
fixes

% The results of this processing on the samples from \cref{tab:data:exectime} can be seen in \cref{tab:data:exectime:processed}.
Three examples of the effect of this discretization and processing on the data are presented in \cref{fig:timing}.
Specifically, we showcase how
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and}}]
    \item higher neuroticism leads to higher mean execution times as impairment goes up (\cref{fig:timing:impneurvsetime})
    \item duration has both a speed-up and slow-down effect on execution times, depending on the level of impairment (\cref{fig:timing:durvsetime})
    \item transitions from higher to lower levels of impairment lead to higher mean execution times when compared to steps far away from a transition, or from a transition from lower to higher levels of impairment (\cref{fig:timing:imptransvsetime})
\end{enumerate*}.

@@

We also implement two simpler reference timing models:
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    % \item a model which returns a constant execution time, corresponding to the mean execution time of all samples associated with the lowest impairment and the shortest duration
    \item a model which randomly samples from the raw data without any grouping
    \item a model which fits an \gls{exGaussian} distribution to the raw data, without any grouping, and then randomly samples from it
\end{enumerate*}.
We will refer to these models individually as the \emph{naive} and \emph{fitted-naive} model, respectively, and collectively as the \emph{firt-order models}.

@@

We show the results of these runs in \cref{fig:ctrlttf}.
Two interesting behaviors of the realistic models can be noticed in this plot.
One, all realistic models are faster than the reference models at the lowest \gls{TTF}.
When compared to the fitted-naive model, the reduction in total duration was on average \textasciitilde\SI{12}{\percent} for realistic models with low neuroticism, and \textasciitilde\SI{6.8}{\percent} for the same models with high neuroticism.

Two, at the highest \gls{TTF}, \emph{only} the high neuroticism models show any significant difference in duration with respect to the reference models.
For \gls{TTF} \(=\) \SI{5.0}{s}, the high neuroticism realistic models were on average \textasciitilde\SI{14.18}{\percent} slower than the fitted-naive model, whereas the low neuroticism models showed no significant relative difference in duration.

>>>>

% The results of this processing on the samples from \cref{tab:data:exectime} can be seen in \cref{tab:data:exectime:processed}.
Three examples of the effect of this discretization and processing on the data are presented in \cref{fig:timing}.
Specifically, we showcase how
\begin{inlineenum}
    \item higher neuroticism leads to higher mean execution times as impairment goes up (\cref{fig:timing:impneurvsetime})
    \item duration has both a speed-up and slow-down effect on execution times, depending on the level of impairment (\cref{fig:timing:durvsetime})
    \item transitions from higher to lower levels of impairment lead to higher mean execution times when compared to steps far away from a transition, or from a transition from lower to higher levels of impairment (\cref{fig:timing:imptransvsetime})
\end{inlineenum}.

@@

We also implement two simpler reference timing models:
\begin{inlineenum}
    % \item a model which returns a constant execution time, corresponding to the mean execution time of all samples associated with the lowest impairment and the shortest duration
    \item a model which randomly samples from the raw data without any grouping
    \item a model which fits an \gls{exGaussian} distribution to the raw data, without any grouping, and then randomly samples from it
\end{inlineenum}.
We will refer to these models individually as the \emph{naive} and \emph{fitted-naive} model, respectively, and collectively as the \emph{firt-order models}.

@@

We show the results of these runs in \cref{fig:ctrlttf}.
Two interesting behaviors of the realistic models can be noticed in this plot.
One, all realistic models are faster than the reference models at the lowest \gls{TTF}.
When compared to the fitted-naive model, the reduction in total duration was on average \ensuremath{\sim}\SI{12}{\percent} for realistic models with low neuroticism, and \ensuremath{\sim}\SI{6.8}{\percent} for the same models with high neuroticism.

Two, at the highest \gls{TTF}, \emph{only} the high neuroticism models show any significant difference in duration with respect to the reference models.
For \gls{TTF} \(=\) \SI{5.0}{s}, the high neuroticism realistic models were on average \ensuremath{\sim}\SI{14.18}{\percent} slower than the fitted-naive model, whereas the low neuroticism models showed no significant relative difference in duration.
